Raft Distributed Consensus Algorithm over\\ WebRTC (Web Real-Time Communication)

### 2015-09-19 research_class/paper.tex
Restore neutral "their" but with plural object.
- But even singular "their" is acceptable:
    http://blog.oxforddictionaries.com/2012/06/he-or-she-versus-they/
Fix typo in previous commit.
Try again to fix typos.
IEEEtran style bibliography.

\documentclass[9pt]{extarticle}

\usepackage{multicol}
\setlength{\columnsep}{0.8cm}

\usepackage{geometry}
\geometry{margin=2.0cm}

%\usepackage[utf8]{inputenc}
\usepackage[backend=bibtex]{biblatex}

\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{authblk}

\usepackage{microtype}

@@

\title{Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)}

\author{Joel D. Martin}
\author{David Levine}
\affil{University of Texas at Arlington}

\addbibresource{references/raft.bib}
\addbibresource{references/w3c.bib}
\addbibresource{references/ietf.bib}
\addbibresource{references/other.bib}

@@

@@

\begin{multicols}{2}

@@

\end{multicols}

\begin{figure*}[h]
\centerline{\includegraphics[width=0.8\textwidth]{imgs/raft_rtc_architecture}}
\caption{Raft over RTC Architecture}
  \label{fig:raft_rtc_architecture}
\end{figure*}

\begin{multicols}{2}

@@

\begin{enumerate}
    \item data transport
    \item data framing and security
    \item connection management
    \item data formats
    \item presentation and control
    \item and local system support functions
\end{enumerate}

@@

The PeerJS client library also provides a default signaling mechanism
that is designed to operate with a PeerJS signaling server. This is
particularly helpful because much of the WebRTC signaling transport
and protocol is not defined as part of the WebRTC standardization
effort. %TODO/CITE
The PeerJS organization provides a cloud-hosted version of a PeerJS
compatible signaling server, however an application developer can also
run her own signaling service using PeerJS PeerServer.

\end{multicols}

\begin{figure*}[h]
\centerline{\includegraphics[width=0.8\textwidth]{imgs/raft_rtc_sequence}}
\caption{Raft over RTC Sequence Diagram}
  \label{fig:raft_rtc_sequence}
\end{figure*}

\begin{multicols}{2}

@@

\begin{center}
    \includegraphics[width=0.45\textwidth]{imgs/browser_1.png}
    \captionof{figure}{Basic app with one Node/Tab}
    \label{fig:browser_1}
\end{center}

@@

\begin{center}
    \includegraphics[width=0.45\textwidth]{imgs/browser_5.png}
    \captionof{figure}{Basic app with Five Nodes/Tabs}
    \label{fig:browser_5}
\end{center}

@@

\begin{center}
    \includegraphics[width=0.45\textwidth]{imgs/chat_1a.png}
    \captionof{figure}{Chat app with one Node/Tab}
    \label{fig:browser_chat_1}
\end{center}

@@

\begin{center}
    \includegraphics[width=0.45\textwidth]{imgs/chat_5a.png}
    \captionof{figure}{Chat app with Five Nodes/Tabs}
    \label{fig:browser_chat_5}
\end{center}

@@

\begin{center}
    \begin{tikzpicture}
    \begin{axis}[%
        xlabel={Nodes}, ylabel={Time (seconds)},
        xtick=data,
        grid=major,
        scatter/classes={%
            propagate={mark=x,draw=blue},
            propagate_average={mark=,draw=red}}]
    \addplot[scatter,only marks,%
        scatter src=explicit symbolic]%
    table[meta=label] {data/data_propagate.dat};
    \addplot[scatter,%
        scatter src=explicit symbolic]%
    table[meta=label] {data/data_propagate_averages.dat};
    \end{axis}
    \end{tikzpicture}
\captionof{figure}{Propagate$,$ Commit$,$ and Apply Command}
\label{fig:propagate}
\end{center}

@@

%\end{multicols}

\begin{figure*}[h]
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et100={mark=x,draw=blue},
                    et100_fail={mark=x,draw=red},
                    et100_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:A}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et500={mark=x,draw=blue},
                    et500_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:B}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et1000={mark=x,draw=blue},
                    et1000_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:C}
    \end{subfigure}
    \caption{Initial Cluster Activation}
    \label{fig:cluster_up}

@@

    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_fail={mark=x,draw=red},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:D}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:E}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:F}
    \end{subfigure}
    \caption{Recovery from Leader Failure}
    \label{fig:recovery-leader-failure}

@@

    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:G}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:H}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:I}
    \end{subfigure}
    \caption{Recovery from Half Cluster Failure}
    \label{fig:recovery-half-failure}
\end{figure*}

%\begin{multicols}{2}

@@

\section*{Acknowledgments}
I would like to thank my academic supervisor, David Levine, for his
excellent input and oversight of this project. I would also like to
thank Diego Ongaro for his amazing work on the Raft protocol and paper
\cite[p.~215]{lamport78} which made implementing
a Raft implementation a pleasure \cite{lamport78}.
There are too many people involved in the creation and standardization
of WebRTC but I would like to thank them for their tireless work to
make browser-to-browser communication possible. Last but not least,
I would like to thank my family for putting up with the long nights
and weekends I spent on this project. Thanks!

\end{multicols}

\printbibliography[title={General references},notkeyword=rtc-general,notkeyword=rtc-transport,notkeyword=rtc-framing]

\nocite{draft-rtcweb-overview}
\nocite{draft-rtcweb-security-arch}
\nocite{draft-rtcweb-security}
\printbibliography[title={WebRTC general references},keyword=rtc-general]

\nocite{draft-rtcweb-transports}
\nocite{draft-rtcweb-qos}
\nocite{RFC4960:sctp}
\nocite{RFC3758:sctp-reliability}
\nocite{draft-sctp-dtls-encaps}
\nocite{draft-mmusic-sctp-sdp}
\nocite{draft-sctp-ndata}
\nocite{draft-rtcweb-rtp-usage}
\nocite{draft-rtcweb-data-channel}
\nocite{draft-rtcweb-data-protocol}
\nocite{draft-rtcweb-alpn}
\nocite{RFC5245:ice}
\nocite{RFC5128:p2p-nat}
\nocite{RFC5389:stun}
\nocite{RFC3489:stun-udp}
\nocite{RFC5766:turn}
\nocite{RFC6156:turn-ipv6}
\nocite{RFC6544:tcp-ice}
\nocite{RFC4571:rtp-rtcp}
\nocite{RFC5764:dtls-srtp}
\printbibliography[title={WebRTC data transport},keyword=rtc-transport]

\nocite{RFC3550:rtp}
\nocite{RFC3711:srtp}
%\nocite{draft-rtcweb-rtp-usage}
%\nocite{draft-rtcweb-data-channel}
%\nocite{draft-rtcweb-data-protocol}
\printbibliography[title={WebRTC framing and security},keyword=rtc-framing]

\nocite{RFC2327:sdp}
\nocite{RFC3264:sdp-offer-answer}
\nocite{draft-rtcweb-jsep}
%\nocite{RFC5245:ice}
\nocite{draft-mmusic-trickle-ice}
\nocite{RFC5763:srtp-using-dtls}
\printbibliography[title={WebRTC connection management},keyword=rtc-connection]

%%%

%\begin{enumerate}
%\item Data formats (audio)
%    \begin{enumerate}
%    \item \href{https://tools.ietf.org/html/draft-cbran-rtcweb-codec-02}{WebRTC: WebRTC Codec and Media Processing Requirements}
%    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-audio-07}{WebRTC: Audio Codec and Processing Requirements}
%    \item \href{https://tools.ietf.org/html/rfc6716}{RFC6716: Opus Audio Codec}
%    \item \href{https://tools.ietf.org/html/draft-ietf-payload-rtp-opus-11}{RTP Payload Format for Opus Speech and Audio Codec}
%    \item \href{https://tools.ietf.org/html/rfc3551}{RFC3551: RTP Profile for Audio and Video Conference}
%    \item \href{https://tools.ietf.org/html/rfc4733}{RFC4733: RTP Payload for DTMF Digits, Telephony Tones, and Telephony Signals}
%    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-audio-codecs-for-interop-01}{WebRTC: Additional WebRTC audio codecs for interoperability with legacy networks}
%    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-video-05}{WebRTC: Video Processing and Codec Requirements}
%    \item \href{https://tools.ietf.org/html/rfc6386}{RFC6386: VP8 Data Format and Decoding Guide}
%    \item \href{http://www.itu.int/rec/T-REC-H.264}{ITU H.264: Advanced video coding for generic audiovisual services}
%    \item \href{https://tools.ietf.org/html/draft-ietf-payload-vp8-15}{RTP Payload Format for VP8 Video}
%    \item \href{https://tools.ietf.org/html/rfc6236}{RFC6236: Negotiation of Generic Image Attributes in the Session Description Protocol (SDP)}
%    \end{enumerate}
%\end{enumerate}

>>>>

\documentclass[conference,compsoc]{./IEEEtran/IEEEtran}

% *** CITATION PACKAGES ***
% IEEE Computer Society needs nocompress option
% requires cite.sty v4.0 or later (November 2003)
\usepackage[nocompress]{cite}

% *** ALIGNMENT PACKAGES ***
\usepackage{array}

% *** SUBFIGURE PACKAGES ***
\usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

%%\usepackage{geometry}
%%\geometry{margin=2.0cm}
%%
%%\usepackage[backend=bibtex]{biblatex}
%%
\usepackage{hyperref}
%%\usepackage{caption}
%%\usepackage{subcaption}
%%
%%\usepackage{authblk}
%%
%%\usepackage{microtype}

@@

%%\title{Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)}
%%
%%\author{Joel D. Martin}
%%\author{David Levine}
%%\affil{University of Texas at Arlington}
%%
%%\addbibresource{references/raft.bib}
%%\addbibresource{references/w3c.bib}
%%\addbibresource{references/ietf.bib}
%%\addbibresource{references/other.bib}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

\title{Raft Distributed Consensus Algorithm over\\ WebRTC (Web Real-Time Communication)}

\author{\IEEEauthorblockN{Joel Martin}
\IEEEauthorblockA{Department of Computer Science and Engineering\\
University of Texas at Arlington\\
Arlington, TX 76010\\
Email: joel@martintribe.org}
\and
\IEEEauthorblockN{David Levine}
\IEEEauthorblockA{Department of Computer Science and Engineering\\
University of Texas at Arlington\\
Arlington, TX 76010\\
Email: levine@cse.uta.edu}}

@@

@@

\begin{figure*}[!t]
\centerline{\includegraphics[width=0.8\textwidth]{imgs/raft_rtc_architecture}}
\caption{Raft over RTC Architecture}
  \label{fig:raft_rtc_architecture}
\end{figure*}

@@

\begin{enumerate}
    \item data transport
    \item data framing and security
    \item connection management
    \item data formats
    \item presentation and control
    \item local system support functions
\end{enumerate}

@@

The PeerJS client library also provides a default signaling mechanism
that is designed to operate with a PeerJS signaling server. This is
particularly helpful because much of the WebRTC signaling transport
and protocol is not defined as part of the WebRTC standardization
effort. %TODO/CITE
The PeerJS organization provides a cloud-hosted version of a PeerJS
compatible signaling server, however application developers can also
run their own signaling service using PeerJS PeerServer.

%\begin{figure*}[!t]
\begin{figure}[!t]
%\centerline{\includegraphics[width=0.8\textwidth]{imgs/raft_rtc_sequence}}
\centerline{\includegraphics[width=0.45\textwidth]{imgs/raft_rtc_sequence}}
\caption{Raft over RTC Sequence Diagram}
  \label{fig:raft_rtc_sequence}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.45\textwidth]{imgs/browser_1.png}
    %\captionof{figure}{Basic app with one Node/Tab}
    \caption{Basic app with one Node/Tab}
    \label{fig:browser_1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.45\textwidth]{imgs/browser_5.png}
    %\captionof{figure}{Basic app with Five Nodes/Tabs}
    \caption{Basic app with Five Nodes/Tabs}
    \label{fig:browser_5}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.45\textwidth]{imgs/chat_1a.png}
    %\captionof{figure}{Chat app with one Node/Tab}
    \caption{Chat app with one Node/Tab}
    \label{fig:browser_chat_1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.45\textwidth]{imgs/chat_5a.png}
    %\captionof{figure}{Chat app with Five Nodes/Tabs}
    \caption{Chat app with Five Nodes/Tabs}
    \label{fig:browser_chat_5}
\end{figure}

@@

\begin{figure}[!t]
    \begin{tikzpicture}
    \begin{axis}[%
        xlabel={Nodes}, ylabel={Time (seconds)},
        xtick=data,
        grid=major,
        scatter/classes={%
            propagate={mark=x,draw=blue},
            propagate_average={mark=,draw=red}}]
    \addplot[scatter,only marks,%
        scatter src=explicit symbolic]%
    table[meta=label] {data/data_propagate.dat};
    \addplot[scatter,%
        scatter src=explicit symbolic]%
    table[meta=label] {data/data_propagate_averages.dat};
    \end{axis}
    \end{tikzpicture}
\caption{Propagate$,$ Commit$,$ and Apply Command}
  \label{fig:propagate}
\end{figure}

@@

%\begin{figure*}[!t]
\begin{figure*}[h]
    \centering
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[100ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et100={mark=x,draw=blue},
                    et100_fail={mark=x,draw=red},
                    et100_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:A}
    }
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[500ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et500={mark=x,draw=blue},
                    et500_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:B}
    }
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[1000ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et1000={mark=x,draw=blue},
                    et1000_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:C}
    }
    \caption{Initial Cluster Activation}
    \label{fig:cluster_up}

@@

    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[100ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_fail={mark=x,draw=red},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:D}
    }
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[500ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:E}
    }
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[1000ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:F}
    }
    \caption{Recovery from Leader Failure}
    \label{fig:recovery-leader-failure}

@@

    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[100ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:G}
    }
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[500ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:H}
    }
    %\begin{subfigure}[b]{0.30\textwidth}
    \subfloat[1000ms election timer]{
        \centering
        \resizebox{0.3\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                %height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \label{fig:I}
    }
    \caption{Recovery from Half Cluster Failure}
    \label{fig:recovery-half-failure}
\end{figure*}

@@

\section*{Acknowledgments}
I would like to thank my academic supervisor, David Levine, for his
excellent input and oversight of this project. I would also like to
thank Diego Ongaro for his amazing work on the Raft protocol and paper which
made implementing a Raft implementation a pleasure.  There are too many people
involved in the creation and standardization of WebRTC but I would like to
thank them for their tireless work to make browser-to-browser communication
possible. Last but not least, I would like to thank my family for putting up
with the long nights and weekends I spent on this project. Thanks!

\bibliographystyle{./IEEEtranBST2/IEEEtran}
\bibliography{./references/ietf,./references/w3c,./references/raft,./references/other}

%% % Bibliography with multiplie sectiosn and references not actually
%% % cited in the paper either

%% \addbibresource{references/raft.bib}
%% \addbibresource{references/w3c.bib}
%% \addbibresource{references/ietf.bib}
%% \addbibresource{references/other.bib}
%%
%% \printbibliography[title={General references},notkeyword=rtc-general,notkeyword=rtc-transport,notkeyword=rtc-framing]
%% 
%% \nocite{draft-rtcweb-overview}
%% \nocite{draft-rtcweb-security-arch}
%% \nocite{draft-rtcweb-security}
%% \printbibliography[title={WebRTC general references},keyword=rtc-general]
%% 
%% \nocite{draft-rtcweb-transports}
%% \nocite{draft-rtcweb-qos}
%% \nocite{RFC4960:sctp}
%% \nocite{RFC3758:sctp-reliability}
%% \nocite{draft-sctp-dtls-encaps}
%% \nocite{draft-mmusic-sctp-sdp}
%% \nocite{draft-sctp-ndata}
%% \nocite{draft-rtcweb-rtp-usage}
%% \nocite{draft-rtcweb-data-channel}
%% \nocite{draft-rtcweb-data-protocol}
%% \nocite{draft-rtcweb-alpn}
%% \nocite{RFC5245:ice}
%% \nocite{RFC5128:p2p-nat}
%% \nocite{RFC5389:stun}
%% \nocite{RFC3489:stun-udp}
%% \nocite{RFC5766:turn}
%% \nocite{RFC6156:turn-ipv6}
%% \nocite{RFC6544:tcp-ice}
%% \nocite{RFC4571:rtp-rtcp}
%% \nocite{RFC5764:dtls-srtp}
%% \printbibliography[title={WebRTC data transport},keyword=rtc-transport]
%% 
%% \nocite{RFC3550:rtp}
%% \nocite{RFC3711:srtp}
%% %\nocite{draft-rtcweb-rtp-usage}
%% %\nocite{draft-rtcweb-data-channel}
%% %\nocite{draft-rtcweb-data-protocol}
%% \printbibliography[title={WebRTC framing and security},keyword=rtc-framing]
%% 
%% \nocite{RFC2327:sdp}
%% \nocite{RFC3264:sdp-offer-answer}
%% \nocite{draft-rtcweb-jsep}
%% %\nocite{RFC5245:ice}
%% \nocite{draft-mmusic-trickle-ice}
%% \nocite{RFC5763:srtp-using-dtls}
%% \printbibliography[title={WebRTC connection management},keyword=rtc-connection]

%% % WebRTC standards that aren't relevant

%% %\begin{enumerate}
%% %\item Data formats (audio)
%% %    \begin{enumerate}
%% %    \item \href{https://tools.ietf.org/html/draft-cbran-rtcweb-codec-02}{WebRTC: WebRTC Codec and Media Processing Requirements}
%% %    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-audio-07}{WebRTC: Audio Codec and Processing Requirements}
%% %    \item \href{https://tools.ietf.org/html/rfc6716}{RFC6716: Opus Audio Codec}
%% %    \item \href{https://tools.ietf.org/html/draft-ietf-payload-rtp-opus-11}{RTP Payload Format for Opus Speech and Audio Codec}
%% %    \item \href{https://tools.ietf.org/html/rfc3551}{RFC3551: RTP Profile for Audio and Video Conference}
%% %    \item \href{https://tools.ietf.org/html/rfc4733}{RFC4733: RTP Payload for DTMF Digits, Telephony Tones, and Telephony Signals}
%% %    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-audio-codecs-for-interop-01}{WebRTC: Additional WebRTC audio codecs for interoperability with legacy networks}
%% %    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-video-05}{WebRTC: Video Processing and Codec Requirements}
%% %    \item \href{https://tools.ietf.org/html/rfc6386}{RFC6386: VP8 Data Format and Decoding Guide}
%% %    \item \href{http://www.itu.int/rec/T-REC-H.264}{ITU H.264: Advanced video coding for generic audiovisual services}
%% %    \item \href{https://tools.ietf.org/html/draft-ietf-payload-vp8-15}{RTP Payload Format for VP8 Video}
%% %    \item \href{https://tools.ietf.org/html/rfc6236}{RFC6236: Negotiation of Generic Image Attributes in the Session Description Protocol (SDP)}
%% %    \end{enumerate}
%% %\end{enumerate}
